---
title: "Chess Tournament Project 1"
author: "Sabina Baraili"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(dplyr)
```

# Project 1: Chess Tournament Cross Table

In this project, we parse the chess tournament cross table and generate a dataset with the following fields:

- **Name**
- **State**
- **Total Points (Score)**
- **Pre-Rating**
- **Average Pre-Rating of Opponents**

We use the raw text file provided and process it step by step.

---

## Step 1: Read Data
```{r}
# Read directly from GitHub raw file (or replace with local path if needed)
url <- "https://raw.githubusercontent.com/savibaraili/data-607-project1/refs/heads/main/tournamentinfo.txt"
lines <- suppressWarnings(readLines(url))

# Separate player rows (names + scores) and state/rating rows
player_rows <- lines[str_detect(lines, "^[[:space:]]*[0-9]+[[:space:]]*\\\\|")]
state_rows  <- lines[str_detect(lines, "^[[:space:]]*[A-Z]{2}[[:space:]]*\\\\|")]

```

---

## Step 2: Extract Player Info
Each player has two rows in the text file:
1. Player row: contains name, total points, and opponents.
2. State row: contains state and pre/post rating.

We extract these pieces into a structured table.
```{r}
players <- tibble(
  Name = str_trim(str_sub(player_rows, 5, 36)),
  Score = as.numeric(str_extract(player_rows, "\\d\\.\\d")),
  State = str_extract(state_rows, "^[A-Z]{2}"),
  PreRating = as.numeric(str_extract(state_rows, "(?<=R: )\\d+")),
  Opponents = str_extract_all(player_rows, "(?<=W |L |D )\\d+")
)
```

---

## Step 3: Calculate Average Opponent Rating
We map each opponent number to its pre-rating, then take the average.
```{r}
get_avg_opp_rating <- function(opps, ratings) {
  opp_ids <- as.numeric(opps)
  mean(ratings[opp_ids], na.rm = TRUE)
}

players <- players %>%
  rowwise() %>%
  mutate(AvgOppRating = round(get_avg_opp_rating(Opponents, PreRating), 0)) %>%
  ungroup()
```

---

## Step 4: Worked Example (Gary Hua)
Let’s confirm the calculation for **Gary Hua** (first player).

```{r}
example <- players[1, ]

example_name <- example$Name
example_state <- example$State
example_score <- example$Score
example_prerating <- example$PreRating
example_opps <- unlist(example$Opponents)
example_opp_ratings <- players$PreRating[as.numeric(example_opps)]
example_avg <- mean(example_opp_ratings)

list(
  Name = example_name,
  State = example_state,
  Score = example_score,
  PreRating = example_prerating,
  Opponents = example_opps,
  OpponentRatings = example_opp_ratings,
  AverageOpponentRating = example_avg
)
```

---

## Step 5: Final Output
We now produce the final dataset.
```{r}
final_df <- players %>%
  select(Name, State, Score, PreRating, AvgOppRating)

# Show only the first 10 rows in the knitted report
head(final_df, 10)
```

---

## Step 6: Save as CSV
All 64 players are saved into a CSV file.
```{r}
write.csv(final_df, "chess_players.csv", row.names = FALSE)
```

**Note:** The CSV contains all rows; this PDF shows only the first 10 for readability.
